---
title: "Workbook 2"
author: "Paul Edwards"
date: "03/03/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fun and figures with R

1.
```{r x1}
#Mathematical calculations
1+1

2-2

3*3

4/4

(5^5)^5

(6^6)^-6

7^-7

8.8*8.8

99+(9/9)

go_factorial <- function(x) {
  y <- 1
  for (i in 1:x) {
    y <- y * ((1:x)[i])
  }
  print(y)
}
go_factorial(10)
```

2.
```{r x2}
#Square root function
sqrt(324)
```

3.
```{r x3}
#Find row with max sepal length and subset dataframe
maxSL <- which.max(iris$Sepal.Length)
iris[maxSL,]
```

4.
```{r x4}
#Find row with min sepal length and subset dataframe
minSL <- which.min(iris$Sepal.Length)
iris[minSL,]
```

5.
```{r x5}
#Calculate range between maximum and minimum sepal length
rangeSL <- max(iris$Sepal.Length)-min(iris$Sepal.Length)
rangeSL
```

6.
```{r x6}
#Construct dataframe with 7 columns and 100 randomly generated rows
df1 <- data.frame(
  a = runif(100),
  b = runif(100),
  c = runif(100),
  d = runif(100),
  e = runif(100),
  f = runif(100),
  g = runif(100)
)
str(df1)
```

7.
```{r x7}
#Copy dataframe and rename columns
df2 <- df1
names(df2)[] <- c("Doc","Grumpy","Happy","Sleepy","Bashful","Sneezy","Dopey")
head(df2)
```

8.
```{r x8}
#Run linear regression model of height by weight
mod1 <- lm(height ~ weight, data = women)
summary(mod1)
```

9.
```{r x9}
#Display model results in a table
sjPlot::tab_model(mod1)
#Display model results in a coefficient plot
sjPlot::plot_model(mod1)
```

10.
```{r x10}
#Display model results in a prediction plot with observations
plot1 <- ggeffects::ggpredict(mod1, terms = "weight")
plot(plot1,
        add.data = TRUE,
        dot.alpha = .8,
        jitter = .2)
```

11.
```{r x11}
#Example calculation
sum(women$weight > 140) / length(women$weight)
```

This calculation is taking the number of women who weigh over 140 pounds, divided by the total number of women in the sample i.e. the proportion of women who weigh over 140 pounds.

12.
```{r x12}
#Calculate proportion of women over the average weight
mwgt <- mean(women$weight)
sum(women$weight > mwgt) / length(women$weight)
```
46.67% of women weigh over the mean weight of women in this sample.

13.
```{r}
#Set plot area to row of 3 graphs
par(mfrow=c(1,3))
#Plot histograms with varying break length
hist(iris$Petal.Length,
    main="Histogram of Iris Petal Length",
    xlab="Petal Length")
hist(iris$Petal.Length,
    main="Histogram of Iris Petal Length",
    xlab="Petal Length",
    breaks=seq(1,7,2))
hist(iris$Petal.Length,
    main="Histogram of Iris Petal Length",
    xlab="Petal Length",
    breaks=seq(1,7,0.2))
```

Histogram breaks determine the size and number of bins, or columns in which data is displayed. Too few bins and interesting points within the distribution will be lost; too many bins will result in noisy, meaningless data.

14.
```{r x14}
#Identify coding error
mh <- mean(women$height)
sum(women$weight > mh) / length(women$height)
```
This code has the height variable mis-typed. It is attempting to sum the weight of women with weights greater than the mean height. The two different scales mean that all women have weights greater than the mean height.

15.
```{r x15}
#Transpose and rename dataframe columns
women2 <- data.frame(women[,c(2,1)])
names(women2)[] <- c("w","h")
str(women2)
```

16.
```{r x16}
#Read data
library(readr)
testdata <- readr::read_csv(url("https://raw.githubusercontent.com/go-bayes/psych-447/main/data/testdata1.csv"))
str(testdata)

#Save data
library(here)
saveRDS(testdata, here::here("data", "td.RDS"))

td <- readRDS(here::here("data", "td.RDS"))
str(td)

#Run linear regression model of height by weight
mod2 <- lm(height ~ weight, data = td)
summary(mod2)

#Display model results in a coefficient plot
sjPlot::plot_model(mod2)

#Display model results in a prediction plot with observations
plot2 <- ggeffects::ggpredict(mod2, terms = "weight")
plot(plot2,
        add.data = TRUE,
        dot.alpha = .8,
        jitter = .2)
```

The intercept represents the predicted value of height when weight is zero. Here, the intercept is nonsensical as it would mean a height of 12.7cm given a weight of 0kg, which is generally not possible unless you are in space.